---
title: "Get more out of your data"
---
# A practical workshop on accessing, processing and visualising data using R programming language

---

First, please install some necessary packages for this tutorial:
tidyverse, httr, jsonlite and plotly. You can install these packages by going to Tools -> Install packages. Here, please select "Install from: Repository (CRAN)" and search for each package separately. Please remember to make sure to "check" the box "Install dependencies".

Links that will be accessed during this workshop:

https://datahelpdesk.worldbank.org/knowledgebase/articles/898581-api-basic-call-structure
https://datahelpdesk.worldbank.org/knowledgebase/articles/898599-api-indicator-queries
https://www.programcreek.com/2014/01/vector-array-list-and-data-frame-in-r/
https://www.json.org/
http://www.r-tutor.com/r-introduction/data-frame
https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf

---
```{r}
# Load packages
library(tidyverse)
library(httr)
library(jsonlite)
library(plyr)

```

```{r}
api <- 'http://api.worldbank.org/v2/' 
```

```{r}
url_list_lic<- 'http://api.worldbank.org/v2/incomeLevels/LIC/countries'
list_lic<-GET(url_list_lic)
list_lic
```

```{r}
list_lic<-GET(url= paste(url_list_lic, '?format=json'))
list_lic
```

```{r}
mrv_url<-'http://api.worldbank.org/v2/countries/chn;bra/indicators/DPANUSIFS?MRV=5'
```

```{r}
mrnev_url<-'http://api.worldbank.org/v2/countries/BR;CR/indicators/SH.STA.MMRT.NE?MRNEV=1'
```

```{r}
gapfill_url_mrv<-'http://api.worldbank.org/v2/countries/chn;bra/indicators/DPANUSIFS?MRV=5&Gapfill=Y'
```

```{r}
freq<-'http://api.worldbank.org/v2/en/countries/ind;chn/indicators/DPANUSSPF?date=2000:2006&MRV=5&frequency=Q'

```

```{r}
get_freq<-GET(freq)
get_freq
```

```{r}
get_json_freq<-GET(url= paste(freq,'?format=json'))
get_json_freq
```

```{r}
list_indicators<-'http://api.worldbank.org/v2/indicators?format=json'
list_indicators_json<-GET(list_indicators)
list_indicators_content<- content(list_indicators_json)
head(list_indicators_content[[1]])
```

```{r}
indicators_list <- fromJSON(list_indicators)
indicators_list
```

```{r}
indicators_list[[2]]$id
```

```{r}
is.list(indicators_list)
is.data.frame(indicators_list)
```

```{r}
indicators_list[[1]]
```

```{r}
indicators_list[[2]]
```

```{r}
# df stands for data frame
indicators_df<- indicators_list[[2]]
```

```{r}
#Let's try if this worked?
is.data.frame(indicators_df)
```

```{r}
names(indicators_df)
```

```{r}
indicators_df[,c("id","name")]
```

```{r}
indicators_list <- lapply(indicators_list, function(x) {
  x[sapply(x, is.null)] <- NA
  unlist(x)
})
```


```{r}
library(plotly)
```

```{r}
a=rnorm(50)
b=rnorm(50)
```

```{r}
help(rnorm)
```


```{r}
p<-plot_ly(x=a, y=b, type='scatter', mode='markers', size=a)
p
```

```{r}

brazil_url<-'http://api.worldbank.org/v2/countries/br/indicators/1.2.PGap.Poor4uds?format=json'

colombia_url<-'http://api.worldbank.org/v2/countries/co/indicators/1.2.PGap.Poor4uds?format=json'

poverty_gap_content_br<-fromJSON(brazil_url)
poverty_gap_content_co<-fromJSON(colombia_url)

##for Colombia the summary page would be
poverty_gap_content_co[[1]]

```

```{r}
##for Brazil the summary page would be
poverty_gap_content_br[[1]]
```
```{r}
brazil_poverty_gap<-poverty_gap_content_br[[2]]
colombia_poverty_gap<-poverty_gap_content_co[[2]]

#What we are interested in here is date and value
colombia_poverty_gap_date_value<-colombia_poverty_gap[,c("date","value")]

brazil_poverty_gap_date_value<-brazil_poverty_gap[,c("date","value")]

```
 
```{r}
colombia_poverty_gap_date_value <- lapply(colombia_poverty_gap_date_value, function(x) {
  x[sapply(x, is.null)] <- NA
  unlist(x)
})

brazil_poverty_gap_date_value <- lapply(brazil_poverty_gap_date_value, function(x) {
  x[sapply(x, is.null)] <- NA
  unlist(x)
})

```

```{r}
brazil<-plot_ly(y=brazil_poverty_gap_date_value$value, x=brazil_poverty_gap_date_value$date, type='scatter', mode='lines')
brazil
```

```{r}
colombia<-plot_ly(y=colombia_poverty_gap_date_value$value, x=colombia_poverty_gap_date_value$date, type='scatter', mode='lines')
colombia
```

```{r}
combined<-plot_ly(y=colombia_poverty_gap_date_value$value, x=colombia_poverty_gap_date_value$date, name='Colombia', type='scatter', mode='lines') %>%
add_trace(y=brazil_poverty_gap_date_value$value, name="Brazil", mode='lines')
combined
```
